---
title: "PSTAT131HW#2"
author: "Praveen Manimaran"
date: "10/16/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(tidymodels)
library(ISLR)
library(ggplot2)
library(corrplot)
tidymodels_prefer()
```
# Exercise 1
Your goal is to predict abalone age, which is calculated as the number of rings plus 1.5. Notice there currently is no `age` variable in the data set. Add `age` to the data set.

Assess and describe the distribution of `age`.

```{r}
abalone <-read.csv('/Users/praveenmanimaran/Desktop/abalone.csv')
abalone$age<-c(abalone$rings+1.5)
abalone

```
Age appears to be randomly distributed with the histogram having approximately having 4 peaks with the data
evenly spread out. 

# Excercise 2
# Split the abalone data into a training set and a testing set. Use stratified sampling. You should decide on appropriate percentages for splitting the data.

*Remember that you'll need to set a seed at the beginning of the document to reproduce your results.*
```{r}

set.seed(3435)

abalone_split <- initial_split(abalone, prop = 0.80,
                                strata = age)
abalone_train <- training(abalone_split)
abalone_test <- testing(abalone_split)


```


### Question 3

Using the **training** data, create a recipe predicting the outcome variable, `age`, with all other predictor variables. Note that you should not include `rings` to predict `age`. Explain why you shouldn't use `rings` to predict `age`.

Steps for your recipe:

1.  dummy code any categorical predictors

2.  create interactions between

    -   `type` and `shucked_weight`,
    -   `longest_shell` and `diameter`,
    -   `shucked_weight` and `shell_weight`

3.  center all predictors, and

4.  scale all predictors.

You'll need to investigate the `tidymodels` documentation to find the appropriate step functions to use.
```{r}

abalone_recipe <-
  recipe(age ~ type+ longest_shell + diameter + height + whole_weight + shucked_weight 
         + viscera_weight +shell_weight, data =   abalone_train)%>% step_dummy(type) %>% 
  step_interact(terms = ~ shucked_weight:starts_with("type")) %>%  
  step_interact(terms = ~ longest_shell:diameter) %>% 
  step_interact(terms = ~ shucked_weight:shell_weight)%>% 
  step_center(all_numeric_predictors()) %>% 
  step_scale(all_numeric_predictors())

```
We shouldn't use rings to predict age because age of an abalone is typically determined by cutting the shell open and counting the number of rings with a microscope. In this assignment we are focused on seeing if we can be accurately predict age using other, easier-to-obtain information about the abalone (excluding rings).


### Question 4

Create and store a linear regression object using the `"lm"` engine.

```{r}

lm_model <- linear_reg() %>% 
  set_engine("lm")

```

### Question 5

Now:

1.  set up an empty workflow,
2.  add the model you created in Question 4, and
3.  add the recipe that you created in Question 3.

```{r}

lm_wflow <- workflow() %>% 
  add_model(lm_model) %>% 
  add_recipe(abalone_recipe)

lm_wflow
```

### Question 6

Use your `fit()` object to predict the age of a hypothetical female abalone with longest_shell = 0.50, diameter = 0.10, height = 0.30, whole_weight = 4, shucked_weight = 1, viscera_weight = 2, shell_weight = 1.

```{r}

lm_fit <- fit(lm_wflow, abalone_train)
lm_fit %>% 
  # This returns the parsnip object:
  extract_fit_parsnip() %>% 
  # Now tidy the linear model object:
  tidy()

abalone_train_res <- predict(lm_fit, new_data = abalone_test %>% select(-age))
abalone_train_res %>% 
  head()

```

### Question 7

Now you want to assess your model's performance. To do this, use the `yardstick` package:

1.  Create a metric set that includes *R^2^*, RMSE (root mean squared error), and MAE (mean absolute error).
2.  Use `predict()` and `bind_cols()` to create a tibble of your model's predicted values from the **training data** along with the actual observed ages (these are needed to assess your model's performance).
3.  Finally, apply your metric set to the tibble, report the results, and interpret the *R^2^* value.


```{r}

#rmse(abalone_train_res, truth = price, estimate = .pred)

abalone_train_res <- predict(lm_fit, new_data = abalone_train %>% select(-age))
abalone_train_res %>% 
  head()

abalone_train_res <- bind_cols(abalone_train_res, abalone_train %>% select(age))
abalone_train_res %>% 
  head()

abalone_metrics <- metric_set(rmse, rsq, mae)
abalone_metrics(abalone_train_res, truth = age, estimate = .pred)


```

The R^2 value was 0.55 which is not close to 1, which means that there is a moderate correlation between the true and predicted values. 
